---
title: "Lab 4: Visualization Redesign and Accessibility"
subtitle: "Student Version"
format:
  html:
    toc: true
    toc-depth: 3
    code-copy: true
    code-tools: true
execute:
  eval: false
  warning: false
  message: false
---

# Learning Objectives

You will critique, redesign, and test a visualization for accessibility.

# Setup

Load `tidyverse` and `gapminder`.

# Part 1: Poor Graph

Using 2007 Gapminder data, create a graph with at least five intentional weaknesses.

```{r}
library(tidyverse)
library(gapminder)
gapminder_2007 <- gapminder %>%
  filter(year == 2007)
```
# Create Graph
```{r}
design_data <- gapminder_2007 %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarize(
    mean_life_expectancy = mean(
      lifeExp,
      na.rm = TRUE
    ),
    .groups = "drop"
  )


ggplot(
  data = design_data,
  mapping = aes(
    x = continent,
    y = mean_life_expectancy,
    fill = continent
  )
) +
  geom_col(
    color = "black",
    linewidth = 1
  ) +
  labs(
    title = "Average Life Expectancy",
    x = "Continent",
    y = "Years",
    fill = "Continent"
  ) +
  theme_gray() +
  theme(
    panel.grid.major = element_line(
      linewidth = 1
    ),
    panel.border = element_rect(
      fill = NA,
      linewidth = 2
    ),
    legend.position = "right"
  )
```



# Part 2: Critique

Identify five problems and explain their effects.

**Response:**  The problems with this graph is their is no year or time period so this does not give clear insight on year or if this is a current graph. The color and legends are reduant because the x- axis is already labeled.Their no sub captions and the graph outline is not needed as well.

# Part 3: Redesign

Create an improved graph with a finding-focused title, meaningful order, accessible design, clear labels, reduced clutter, alt text, and source.

```{r}
ggplot(
  data = design_data,
  mapping = aes(
    x = continent,
    y = mean_life_expectancy
  )
) +
  geom_col(
    linewidth = 1
  ) +
  labs(
    title = "Life Expectancy Varied Substantially by Continent",
    subtitle = "Country-level mean life expectancy in 2007",
    x = NULL,
    y = "Mean life expectancy in years",
    caption = "Source: Gapminder"
  ) +
  
  theme_gray() 

```

# Part 4: Highlighting

Create a graph that highlights one continent and mutes the others.

```{r}

  highlight_data <- design_data %>%
  mutate(
    highlight_group = if_else(
      continent == "Africa",
      "Africa",
      "Other countries"
    )
  )

ggplot(
  data = highlight_data,
  mapping = aes(
    x = continent,
    y = mean_life_expectancy,
    fill = highlight_group
  )
) +
  geom_col() +
  scale_fill_manual(
    values = c("Africa" = "#C2410C", "Other countries" = "gray80")
  ) +
  labs(
    title = "Life Expectancy Varied Substantially by Continent",
    subtitle = "Country-level mean life expectancy in 2007",
    x = NULL,
    y = "Mean life expectancy in years",
    caption = "Source: Gapminder",
    fill = NULL
  )
  

```

# Part 5: Accessibility Review

Evaluate color dependence, contrast, font size, legend use, grayscale readability, and alt text.

**Response:**  The color dependence is quite noctieable, also they major contrast so I can see clearly since continent the graph wants us to draw attention to, the font size is approatrie and the legend is not really needed but still a nice touch. The grayscale readability is also pretty good.

# Part 6: Published Graphic Critique

Critique an instructor-approved public health graphic for purpose, data, chart choice, accessibility, uncertainty, and ethics.

**Response:** I looked at Columbia Law School's Climate Change Law department, which examined COVID-19 data misrepresented by the Georgia Department of Public Health. The purpose of the graph was to show the top five counties in Georgia with the highest numbers of COVID-19 cases. The data itself is okay, but it could be organized better so it's easier to understand. The chart choice is again okay I think their a lot information for a bar chart. The biggest problem is that neither the x-axis nor the y-axis is labeled, and the bars appear to follow no logical order