---
title: "Lab 3: Data Wrangling for Visualization"
subtitle: "Student Version"
format:
  html:
    toc: true
    toc-depth: 3
    code-copy: true
    code-tools: true
execute:
  eval: false
  warning: false
  message: false
---

# Learning Objectives

You will clean, reshape, summarize, and join public health data before visualization.

# Setup

Load `tidyverse`.

```{r}

library(tidyverse)

# Add your code here.
```

# Part 1: Create Raw Data

Create `raw_health_data` with at least eight counties, region, 2023 and 2024 population and case columns, one duplicate county, and one missing value.

```{r}
library(tibble)

raw_health_data <- tibble(
  County     = c("Isle of Wight","York","James City","Southampton","Fairfax",
                 "Loudoun","Prince William","Arlington","Halifax",
                 "Buckingham","Wise","Spotsylvania","Fairfax"),
  region     = c("Hampton Roads","Hampton Roads","Hampton Roads","Hampton Roads",
                 "Northern Virginia","North Virginia","Northern Virginia",
                 "Northern Virginia","Central","Piedmont","Central","South",
                 "Northern Virginia"),
  pop_2023   = c(41200, 73800, 81500, 17600, 1148000,
                 438500, 492300, 236100, 32900,
                 16800, 34700, 146200, 1148000),
  pop_2024   = c(42050, 74600, 83100, 17350, 1157400,
                 449800, 501700, 238900, 32400,
                 16650, 34100, 149800, 1157400),
  cases_2023 = c(312, 486, 574, 198, 8940,
                 2810, 3960, 1520, 405,
                 NA, 521, 1180, 8940),
  cases_2024 = c(298, 512, 601, 215, 9375,
                 3040, 3785, 1490, 388,
                 224, 498, 1265, 9375)
)


# Add your code here.
```


# Part 2: Diagnose

Use `glimpse()`, duplicate checks, and missing-value summaries.

```{r}
glimpse(raw_health_data)
```

# Part 3: Remove Duplicates

Create `health_data_clean` with one row per county.

```{r}

health_data_clean <- raw_health_data %>%
  mutate(region = if_else(region == "North Virginia", "Northern Virginia", region)) %>%
  distinct(County, .keep_all = TRUE)

```

# Part 4: Reshape Long

Create county, region, year, population, and cases columns.

```{r}
health_data_long <- health_data_clean %>%
  pivot_longer(
    cols = c(pop_2023, pop_2024, cases_2023, cases_2024),
    names_to = c(".value", "year"),
    names_sep = "_"
  ) %>%
  rename(county = County, population = pop) %>%
  mutate(year = as.integer(year))
```

# Part 5: Calculate Rates

Calculate cases per 100,000.

```{r}
health_data_long <- health_data_long %>%
  mutate(cases_per_100k = cases / population * 100000)
```

# Part 6: Summarize

Calculate mean rate by region and year.

```{r}
region_rates <- health_data_long %>%
  group_by(region, year) %>%
  summarise(
    mean_rate = mean(cases_per_100k, na.rm = TRUE),
    n_counties = n(),
    .groups = "drop"
  )
```

# Part 7: Join Contextual Data

Create a county rural-status table and join it with `left_join()`. Check duplicates, unmatched rows, and row counts.

```{r}
county_context <- tibble(
  county = c("Isle of Wight","York","James City","Southampton","Fairfax",
             "Loudoun","Prince William","Arlington","Halifax",
             "Buckingham","Wise","Spotsylvania","Accomack"),
  rural_status = c("Rural","Suburban","Suburban","Rural","Urban",
                   "Suburban","Suburban","Urban","Rural",
                   "Rural","Rural","Suburban","Rural"),
  metro_area = c("Virginia Beach-Norfolk","Virginia Beach-Norfolk",
                 "Virginia Beach-Norfolk","Virginia Beach-Norfolk",
                 "Washington DC","Washington DC","Washington DC",
                 "Washington DC",NA,NA,NA,"Washington DC",NA)
)
health_data_joined <- health_data_long %>%
  left_join(county_context, by = "county")
```

# Part 8: Visualize

Create a regional line graph, ranked 2024 county graph, and rural-status comparison.

```{r}
region_rates <- health_data_joined %>%
  group_by(region, year) %>%
  summarise(mean_rate = mean(cases_per_100k, na.rm = TRUE),
            .groups = "drop")

ggplot(region_rates, aes(x = year, y = mean_rate, color = region)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = c(2023, 2024)) +
  labs(title = "Mean case rate by region, 2023-2024",
       x = NULL, y = "Cases per 100,000",
       color = "Region") +
  theme_minimal()
```
```{r}
health_data_joined %>%
  filter(year == 2024) %>%
  ggplot(aes(x = reorder(county, cases_per_100k), y = cases_per_100k,
             fill = rural_status)) +
  geom_col() +
  coord_flip() +
  labs(title = "Case rate by county, 2024",
       x = NULL, y = "Cases per 100,000",
       fill = "Rural status") +
  theme_minimal()
```

```{r}
health_data_joined %>%
  filter(year == 2024) %>%
  ggplot(aes(x = rural_status, y = cases_per_100k, fill = rural_status)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.15, size = 2) +
  labs(title = "Case rate by rural status, 2024",
       x = NULL, y = "Cases per 100,000") +
  theme_minimal() +
  theme(legend.position = "none")
```

# Reflection

Explain why long format was useful, what causes row multiplication during joins, and why rates are preferable to counts.

**Response:**  Long format is useful in ggplot because one columns identifies the grouping variable and the other one looks at the measured value. What causes row multiplication during joins happens because you have the original datasets then it adds matching information from the second datasets.Rates are preferable to counts because it allows you to get a general understanding of population size instead of looking at every specific count which can become time-consuming.
